---
title: "Data visualization"
author: "Jubril"
date: "6/29/2022"
output: html_document
---


Scatter plot

Customization of scatter plots is very much acceptable in R
Examples are:
  Main : The title for the plot ( Displayed at the Top)
  Sub : The subtitle for the plot( displayed at the Bottom)
  Xlim : The X-axis scale; Uses the format c(min,max);
automatically determined by default 
  ylim : The Y-axis: Uses the format c(min,max);
automatically determine by default
  xlab: The x-axis title
  ylab : the Y-axis title
  
  col:Determines the colours used for points and lines;accepts the charracter strings of colour names (red,green,e.t.c)
  pch : the type of point to use ( Circle, Square, Triangle e.t.c); accepts values 0 to 25 for symbols and 32 to 255 for characters
  cex: The amount to scale the size of points; accept a numeric value; default is 1
  lty : defines the line type; accepts various characters strings (solid and dashed, dotted, e.t.c)
  lwd : Defines the line width; accepts a positive number; default is 1
  
```{r}
x<- rnorm(200)
plot(x)
```
  
  Plotting a scatter with a sequence of numbers between -5 and 5, incrementing by 0.1. choose the mean as 2 and standard deviation 0.8

```{r}
x<- seq(-5,5, by = 0.1)
y <- dnorm(x,mean =2, sd = 0.8)

plot(x,y)
```
 The following code will create a scatter plot for the relation between wt(WEIGHT) AND MPG(MILESPERGALLON)  using mtcars dataset
 
```{r}
mtcars
```
 
```{r}
?mtcars
# to chect the info about the data
```

```{r}
data(mtcars)
plot(x = mtcars$wt, y= mtcars$mpg,
     main = "scatterplot", xlab = "car weigth", ylab = "Miles per Gallon")
```
```{r}
plot(x=mtcars$wt,
     y=mtcars$mpg,
     main = "scatterplot",
     xlab = "car weight",
     ylab = "Miles per Gallon",
     xlim = c(2.5,5),
     ylim = c(10,30),
     pch = 8,
     col="red")
```

```{r}
iris
```

```{r}
?iris
```

```{r}
#iris
names(iris)
iris$Species
```
```{r}
unique(iris$Species)
```
```{r}
iris
plot(data=iris,
     main="iris Data",
     Sepal.Length ~ Petal.Length,
     xlab="petal lenght (cm)",
     ylab="Sepal Lenght (cm)",
     pch = c(10,17,18) [as.numeric(Species)],
     col=c("magenta","purple","blue")[as.numeric(Species)])
```
```{r}
x<-rnorm(100)
x
```
```{r}
y<- 3*x + rnorm(100)
y
```
```{r}
z<- sqrt(1+y^2)+ rnorm(100)
z
```

```{r}
x<-rnorm(100)
y<- 3*x + rnorm(100)
z<- sqrt(1+y^2)+ rnorm(100)
plot(x,y,
     main = "conditional scatter plot",
     pch = ifelse(x*y > 5,10,20),
     xlim = range(x),
     ylim = range(y,z),
     xlab = "x",
     ylab = "value",
     col = ifelse(y >= mean(y), "blue","magenta"))
```
Simple Bar chart
```{r}
mtcars
Counts <- table(mtcars$gear)
Counts
```

```{r}
barplot(Counts,main = "Car Distribution by gear",
        xlab ="Number of Gears")
```

```{r}
barplot(Counts,main = "Car distribution by Gear",
        ylab = "Number of Gears",horiz=TRUE)
```

```{r}
mtcars
Counts <- table(mtcars$vs,mtcars$gear)
Counts
```

```{r}
barplot(Counts, main = "Car distribution by gear and vs",
        beside = TRUE,
        horiz = TRUE,
        ylab = "Number of Gears",
        xlab = "Frequency",
        col = c("Lightblue","brown"),
        legend = rownames(Counts))
```

```{r}
colors <- c("green","orange","brown")
Months <- c("mar", "apr","may","jul","aug")
Regions <- c("East","West","North")

values <- matrix(c(2,9,3,11,9,4,8,7,3,12,5,2,8,10,11),
                 nrow = 3, ncol = 5, byrow = TRUE)
 barplot(values,main = "total revenue", names.arg = Months,
         las = 1,xlab = "Months", ylab = "revenue",
         col=colors , beside = TRUE, border = "red", font.axis =7,
         col.axis = "green",cex.axis = 1, col.lab = "red")
 lengend("topleft",Regions,cex = 0.5, fill = colors)
 values
```

```{r}
percentage <- c(10,15,38,28,10)
labels <- c("0-8 yrs 10%","9-11 yrs 15%", "12 yrs 38%","13-15 yrs 28%", "16 and above yrs 10% ")
pie(percentage,labels = labels, main ="Education percent", col = c("blue","yellow","Green","red","brown"))
```
```{r}
X <-c(21,62,10,53)
labels <- c("LONDON","NEW YORK","SINGAPORE","MUMBAI")
PIEPERCENT <- round(100*X/sum(X),1)

pie(X, labels =PIEPERCENT,main ="city pie chart",
    col = rainbow(length(X)))
legend("bottomright", c("LONDON","NEW YORK","SINGAPORE","MUMBAI"), cex = 0.8, fill = rainbow(length(X)))
```
LiNE CHART
```{r}
V <- c(9,12,17,23,21,30,35,31,36)
X <- c(1,2,3,4,5,6,7,8,9)
plot(X,V, type = "o", col = "blue", xlab = "month", ylab = "Growth", main = "Growth Rate")

```
HISTOGRAM
```{r}
random = rnorm(100)
hist(random,border = "Blue",col = "red")
```
```{r}
AirPassengers
hist(AirPassengers, main = "Histogram for air passengers",
     xlab = "passengers",
     border = "blue",
     col = "yellow",
     breaks = 6)
```

```{r}
ToothGrowth
hist(ToothGrowth$len, main = "Histogram for tooth growth",
     xlab = "length",
     border = "blue",
     col = "red",
     breaks = 6)
```
Learning ggplot

ggplot (data = <DATA> ) + GEOM_FUNCTION (mapping = aes( <MAPPINGS>), stat = <STAT> , position = <POSITION> ) + COORDINATE_FUNCTION> + FACET_FUNCTION> +SCALE_FUNCTION> + THEME_FUNCTION>
```{r}
library(ggplot2)
```


```{r}
?ggplot2
```

```{r}
a <- ggplot(economics, aes(date, unemploy))
a + geom_path(lineend="butt", linejoin="round", linemitre=1)
```
```{r}
a + geom_polygon(aes(group = group))
```
```{r}
a + geom_ribbon(aes(ymin=unemploy - 900, ymax=unemploy + 900))
```

```{r}
 b <- ggplot(seals, aes(x = long, y = lat))
b + geom_curve(aes(yend = lat + 1, xend=long+1),curvature=1)
```
```{r}
b + geom_rect(aes(xmin = long, ymin=lat, xmax= long + 1, ymax = lat + 1)) 
```

```{r}
c <- ggplot(mpg,aes(hwy))
c + geom_area(stat="bin")
```

```{r}
c + geom_density(kernel = "gaussian")
```

```{r}
c + geom_dotplot()
```

```{r}
c + geom_freqpoly()
```

```{r}
c + geom_histogram(binwidth = 5)
```

```{r}
e <- ggplot(mpg, aes(cty, hwy))
e + geom_label(aes(label = cty), nudge_x = 1, nudge_y = 1, check_overlap = TRUE) 
```

```{r}
e + geom_jitter(height = 2, width = 2) 
```
```{r}
e + geom_point()
```
```{r}
f <- ggplot(mpg, aes(class, hwy))
f + geom_col()
```
```{r}
f + geom_boxplot()
```
```{r}
d <- ggplot(mpg, aes(fl))
d + geom_bar() 
```
```{r}
n <- d + geom_bar(aes(fill = fl))
n + scale_fill_brewer(palette = "Reds")
```
```{r}
s <- ggplot(mpg, aes(fl, fill = drv))
s + geom_bar(position = "dodge")
```


